# Week 3 Notes


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Subsetting and Ordering

```{r}
x<-data.frame("var1"=sample(1:5), "var2"=sample(6:10), "var3"=sample(11:15))
x<-x[sample(1:5),]; x$var2[c(1,3)]=NA
x
```

-   Subsetting

```{r}
x[,1]
x[,"var1"]
x[1:2,"var2"]
x[(x$var1<=3&x$var3>11),]
x[(x$var1<=3|x$var3>15),]
x[which(x$var2>8),]
```

-   Sorting

```{r}
sort(x$var1)
sort(x$var1,decreasing = TRUE)
sort(x$var2,na.last = TRUE)
```

-   Ordering

```{r}
x[order(x$var1),]
x[order(x$var1,x$var3),]
```

-   Ordering with plyr

```{r}
library(plyr)
arrange(x,var1)
arrange(x,desc(var1))
```

-   Add rows and columns

```{r}
x$var4<-rnorm(5)
x
y<-cbind(rnorm(5),x)
y
```

## Summarizing Data

reading in data file
```{r}
restData<-read.csv("Restaurants.csv")
```

- Look at Data
```{r}
head(restData,n=3)
tail(restData,n=3)
summary(restData)
str(restData)
```

- Summarize Data
```{r}
quantile(restData$cncldst,na.rm = TRUE)
quantile(restData$cncldst,probs = c(0.5,0.75,0.9))
# useNA will tell you if there are missing values, without it it won't report that number
table(restData$zipcode,useNA = "ifany")
table(restData$cncldst,restData$zipcode)
```

- Missing values
```{r}
sum(is.na(restData$cncldst))
any(is.na(restData$cncldst))
all(restData$zipcode>0)
```

- Row and Column Sums
```{r}
colSums(is.na(restData))
all(colSums(is.na(restData))==0)
```

- Specific characteristics
```{r}
table(restData$zipcode %in% c("21212"))
table(restData$zipcode %in% c("21212","21213"))
restData[restData$zipcode %in% c("21212","21213"),]
```

- Cross Tabs
```{r}
data("UCBAdmissions")
DF=as.data.frame(UCBAdmissions)
summary(DF)

xt<-xtabs(Freq~Gender+Admit,data = DF)
xt

warpbreaks$replicate<-rep(1:9,len=54)
yt = xtabs(breaks~.,data = warpbreaks)
yt
```

- Flat Tables
```{r}
ftable(xt)
```

- Size of data set
```{r}
fakeData=rnorm(1e5)
object.size(fakeData)

print(object.size(fakeData),units="Mb")
```

## Creating New Variables

- Creating Sequences
```{r}
s1<-seq(1,10,by=2);s1
s2<-seq(1,10,length=3);s2
x<-c(1,3,8,25,100); seq(along=x)
```

- Subsetting Variables
```{r}
restData$nearMe = restData$nghbrhd %in% c("Roland Park","Homeland")
table(restData$nearMe)
```

- Binary Variables
```{r}
restData$zipWrong = ifelse(restData$zipcode<0, TRUE, FALSE)
table(restData$zipWrong,restData$zipcode<0)
```

- Creating categorical variables
```{r}
restData$zipGroups = cut(restData$zipcode,breaks = quantile(restData$zipcode))
table(restData$zipGroups)
```

- Easier Cutting
```{r}
library(Hmisc)
restData$zipGroups = cut2(restData$zipcode,g=4)
table(restData$zipGroups)
```

- Create Factor Variables
```{r}
restData$zcf<-factor(restData$zipcode)
restData$zcf[1:10]
class(restData$zcf)

yesno<-sample(c("yes","no"),size = 10,replace = TRUE)
yesnofac<-factor(yesno,levels = c("yes","no"))
relevel(yesnofac,ref = "yes")
as.numeric(yesnofac)
```















